Robust Control of a Fluxonium Qubit

### 2020-11-25 paper/0abstract.tex
brian edits

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zac Manchester}
\affiliation{
  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity gates on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control techniques have proven effective in experimentally
  realizing high fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to supress gate errors arising from system
  parameter deviations and noise.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentation.
  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
  integrating the Lindblad master equation, which is computationally expensive.
  We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We demonstrate these techniques numerically on a fluxonium qubit with realisitc
  experimental parameters and constraints,
  achieving orders of magnitude gate error reductions from our baseline gate set.
\end{abstract}

>>>>

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zachary Manchester}
\affiliation{
  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
  % Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high-fidelity gates on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control techniques have proven effective in experimentally
  realizing high-fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  parameter deviations and noise.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentiation.
  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
  integrating the Lindblad master equation, which is computationally expensive.
  We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We demonstrate these techniques numerically on a fluxonium qubit with realistic
  experimental parameters and constraints,
  achieving orders of magnitude gate error reductions from our baseline gate set.
\end{abstract}

### 2020-11-25 paper/1introduction.tex

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}. For quantum computation
QOC techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.
Experimental errors may cause the system evolution to deviate from that predicted in
optimization, hampering performance.
The field of classical control theory has developed robust control techniques
to encode experimental errors in optimization objectives, improving
experimental performance (citations appreciated).
In this work we employ robust control techniques to mitigate
realistic system parameter deviations and decoherence that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate
errors due to parameter deviations.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
longitudinal relaxation and pure dephasing decoherence
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work we mitigate
errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
on the fluxonium qubit \cite{zhang2020universal} by solving a robust trajectory 
optimization problem. 
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve this optimization problem using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}. 
% The augmented Lagrangian method
% allows us to enforce arbitrary, simultaneous constraints and achieve
% fast convergence without being restricted to the constraint
% manifold. 
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
encode the dependence of longitudinal relaxation
on the controls in an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Additionally, to make the quantum state trajectory robust to parameter deviations,
we consider three different approaches to making the optimized trajectory less 
sensitive to model uncertainty:
% we consider three robust
% robust trajectory opitmization techniques to make the
% quantum state trajectory robust to parameter deviations.
\begin{enumerate}
\item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling,
  rembold2020introduction}.
  \item An unscented sampling method, adapted from the unscented tranform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
  to compute derivative information of parameter deviations with respect to the 
  quantum state trajectory.
\end{enumerate}
% The first is the sampling method, which has been applied
% previously in the context of QOC
%%TODO: the citations for unscented sampling
%% and sampling are the same as in section V
% \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
% The second is the unscented sampling
% method, which derives from the unscented transformation used
% for non-linear Kalman filtering
% \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
% We propose a third method, which uses derivative information
% of parameter deviations with respect to the quantum state trajectory.
% We make this method efficient by employing mixed-mode differentiation.

This paper is organized as follows.
First, we introduce the ALTRO method in the context of QOC \ref{sec:background}.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method \ref{sec:fluxonium}. Then, we
outline a method for making the optimization aware of longitudinal
relaxation \ref{sec:longitude}. Next, we outline three methods for achieving
robustness to static parameter deviations \ref{sec:static}. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise
\ref{sec:stochastic}.

>>>>

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
For quantum computation
QOC techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system's trajectory to deviate from that predicted in
optimization, hampering performance.
Robust control techniques 
address this issue by encoding
experimental errors in optimization objectives, yielding performance
garauntees in the presence of model uncertainty \cite{Zhou97,Morimoto00,Manchester18}.
In this work we employ robust control techniques to mitigate
realistic system parameter deviations and decoherence that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates \todo{citations needed}. 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
In order to address these issues,
recent work has sought to mitigate
errors due to parameter deviations.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
longitudinal relaxation and pure dephasing decoherence
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work, we study three robust control techniques that
make the optimized trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
\item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling,
  rembold2020introduction}.
  \item An unscented sampling method, adapted from the unscented tranform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
  to compute derivative information of parameter deviations with respect to the 
  quantum state trajectory.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

This paper is organized as follows.
First, we introduce the ALTRO method in the context of QOC
in Section \ref{sec:background}.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method in Section \ref{sec:fluxonium}. Then, we
outline a method for making the optimization aware of longitudinal
relaxation in Section \ref{sec:longitude}. Next, we outline three methods for achieving
robustness to static parameter deviations in Section \ref{sec:static}. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise
in Section \ref{sec:stochastic}.

### 2020-11-25 paper/2background.tex

\section{QOC + AL-iLQR}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE)
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{clarkson2010coresets, hauswirth2016projected}.

@@

ALM converges superlinearly but tends to exhibit slow constraint convergence near the optimal 
solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
solution-polishing phase that takes 1-2 Newton steps on the active constraint set to provide 
machine-precision constraint satisfaction. For more information on the details of the ALTRO
solver, see \cite{howell2019altro,Jackson2020altroc}.

% For the QOC problem, the state $x$ consists of \todo{fill this out} while the control input
% $u$ includes the flux drive amplitude. The cost functions $\ell_N$
% The trajectory optimization
% literature, driven primarily by the robotics community,
% solves a more general class of non-linear programs that resemble
% the QOC problem. The QOC
% problem is a specific case of the linear quadratic regulator (LQR).
% LQR is so called because the dynamics are linear in the state and
% the functional is quadratic in the state. A general LQR cost
% function takes the form
% \begin{equation}
%   J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
%   + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
% \end{equation}
% The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
% and $u$ is the augmented control vector. The augmented state is propagated
% to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
% In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
% and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
% cost function $\ell_{k}$ includes contributions from the infidelity
% $1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
% $\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
% and control regularization $u^{T}_{k} R_{k} u_{k}$.
% Placing cost functions at each knot point smoothens the optimization
% landscape. In QOC realizing the final state
% is the most important objective so the cost function at the final knot point
% carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
% The LQR formulation restricts the cost functions to operate on the augmented
% state and controls at each knot point. The advantage of the LQR formulation
% is that there exists a differentiable dynamic programming algorithm to compute the
% optimal update to the augmented controls $u_{k}$ which minimizes the cost
% function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
% recurrence relation between knot points $k$ and $k + 1$ for the optimal
% feedback law--known as the Ricatti recursion (citations needed). The
% iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
% and applies the Ricatti recursion to all knot points on multiple
% executions.

% In order to incorporate constraints we employ
% the augmented Lagrangian method. Constraints are contributions
% to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
% zero or negative when the constraint is satisfied. The AL-iLQR
% method associates a penalty multiplier with the constraint
% that estimates the constraint's Lagrange multiplier.
% The algorithm updates the penalty multiplier between
% iLQR executions. In this scheme the functional takes the form
% \begin{equation}
% \begin{aligned}
%     J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
%     &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
%     &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
% \end{aligned}
% \end{equation}
% Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
% with $\mu_{k}$ along the diagonal.
% $\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
% \begin{align}
%   \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
%   \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
% \end{align}
% Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
% $i$ indicates the $i$-th constraint functional,
% and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
% recursion, can be formed to calculate the optimal control updates for this functional,
% see \cite{howell2019altro}.

>>>>

\section{QOC + AL-iLQR \label{sec:background}}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
In general $J$ is a linear combinaion of cost functions on the state
as well as cost functions on the controls \cite{leung2017speedup}.
Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{clarkson2010coresets, hauswirth2016projected}.

@@

ALM converges superlinearly but tends to exhibit slow constraint convergence near the optimal 
solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
solution-polishing phase that takes 1-2 Newton steps on the active constraint set to provide 
machine-precision constraint satisfaction. For more information on the details of the ALTRO
solver, see \cite{howell2019altro, Jackson2020altroc}.

The ALTRO method has fast tail-convergence and
allows us to impose arbitrary constraints without restricting
the optimization to the constraint manifold.
This contrasts typical first-order QOC methods that employ the same
update procedure for the duration of optimization
and employ projected gradient methods
which restrict the optimization to the constraint manifold.
To see how to use the ALTRO method for QOC, we will present
a concrete example in the following section.

### 2020-11-25 paper/3fluxonium.tex

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{x_1}{= x_\text{init}} \label{eq:init_con}
    \addConstraint{x_N}{= x_f} \label{eq:term_con}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
\end{equation}
The initial and final conditions are determined by the
desired unitary transformation, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
\end{equation}
The dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
\eqref{eq:tdse}, \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.
The final state transfer is the most important objective which is
encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.
%%, $k \in \{1, ..., N - 1\}$.

Additionally, we impose constraints that reflect limitations
of the apparatus and ones that improve the experimental realization
of the control pulse. To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$.

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude \ref{sec:longitude} and employ robust control techniques to mitigate
pure dephasing \ref{sec:static}, \ref{sec:stochastic}.

>>>>

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1, x_N \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial and final conditions are determined by the
desired unitary transformation, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
\end{equation}
The discrete dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
\eqref{eq:tdse}, \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.
The final state transfer is the most important objective which is
encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$.

%% In addition to obeying experimental constraints
%% and achieving low simulated gate errors, we desire
%% to make our gates perform well in the presence of decoherence.
%% Decoherence of the quantum state due to external noise
%% is modeled by two phenomena: longitudinal relaxation and pure dephasing.
%% They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
%% Dissipation to the thermal bath via longitudinal
%% relaxation is an irreversible process
%% that results in information loss.
%% Converesely, pure dephasing is a reversible process.
%% We choose to maximize the longitudinal relaxation time directly
%% by tuning the flux amplitude, as in Section \ref{sec:longitude}, and
%% employ robust control techniques to mitigate
%% pure dephasing in Section \ref{sec:stochastic}.

%% The main contributions to longitudinal relaxation in our
%% device are dielectric loss in the capacitor, resistive loss in the inductor,
%% and Purcell loss. The main contributions to pure dephasing in our
%% device are $1/f$ flux noise and decay via charge and flux coupling
%% to the control lines.
%% Our device achieves its best pure dephasing
%% protection at the flux frustration point
%% $T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
%% where the qubit is first-order insensitive to changes in flux.
%% It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
%% frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
%% at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
%% and increases away from the flux frustration point
%% $T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
%% of the decay processes and the tradeoff,
